---
title: "p8105_hw6_sc5558"
author: "Shiyun Angel Cheng"
date: "2025-12-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Import Necessary Libraries
```{r}
library(tidyverse)
library(broom)
library(p8105.datasets)
```

## Problem 1
```{r}
#Load and Clean Data
homicides = read_csv("homicide-data.csv") |>
  mutate(
    city_state = str_c(city, ", ", state),
    solved = case_when(
      disposition == "Closed by arrest" ~ 1,
      TRUE ~ 0
    ),
    victim_age = as.numeric(victim_age)
  ) |>
  filter(
    !(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO")),
    city_state != "Tulsa, AL",
    victim_race %in% c("White", "Black")
  )
```
Logistic Regression - Baltimore, MD
```{r}
baltimore_df = homicides |> filter(city_state == "Baltimore, MD")

baltimore_fit = glm(
  solved ~ victim_age + victim_sex + victim_race,
  family = binomial,
  data = baltimore_df
)

baltimore_results =
  tidy(baltimore_fit, conf.int = TRUE, exponentiate = TRUE)

baltimore_results
```
Extract adjusted OR
```{r}
baltimore_or =
  baltimore_results |> 
  filter(term == "victim_sexMale") |> 
  select(estimate, conf.low, conf.high)
```

Fit Models for ALL Cities
```{r}
city_results = 
  homicides |>
  group_by(city_state) |>
  nest() |>
  mutate(
    model = map(data, ~ glm(
      solved ~ victim_age + victim_sex + victim_race,
      family = binomial,
      data = .x
    )),
    tidy_model = map(model, ~ tidy(.x, conf.int = TRUE, exponentiate = TRUE))
  ) |>
  unnest(tidy_model) |>
  filter(term == "victim_sexMale")
```
Plot Adjusted ORs by City 
```{r}
city_results |>
  mutate(city_state = fct_reorder(city_state, estimate)) |>
  ggplot(aes(x = estimate, y = city_state)) +
  geom_point() +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(
    x = "Adjusted Odds Ratio (male vs female)",
    y = "City",
    title = "Adjusted OR for Male vs Female Victims — Homicide Solved Probability"
  )
```
Text Comment: 
Cities vary widely in the adjusted odds of a homicide being solved for male versus female victims. Several cities have OR estimates above 1, indicating that homicides of male victims are more likely to be solved relative to female victims, while others have ORs below 1, showing the opposite pattern. Confidence intervals are often wide, reflecting small sample sizes. Cities with narrower CIs tend to have larger homicide counts and more stable estimates.
## Problem 2

Load Data
```{r}
data("weather_df")

weather_df = 
  weather_df |>
  drop_na(tmax, tmin, prcp)
```
Fit the regression we're boostrapping around: 
```{r}
orig_fit = lm(tmax ~ tmin + prcp, data = weather_df)

glance(orig_fit)    # contains r.squared
tidy(orig_fit)      # contains coefficients, including beta1 / beta2
```
Define a function to do one bootstrap replicate
```{r}
boot_once = function(df) {
  
  boot_sample = df |> 
    slice_sample(prop = 1, replace = TRUE)
  
  fit = lm(tmax ~ tmin + prcp, data = boot_sample)
  
  g = glance(fit)
  t = tidy(fit)
  
  r2 = g$r.squared
  
  beta1 = t$estimate[t$term == "tmin"]
  beta2 = t$estimate[t$term == "prcp"]
  
  tibble(
    r2 = r2,
    beta_ratio = beta1 / beta2
  )
}
```
Run 5000 bootstrap samples
```{r}
set.seed(8105)

boot_results = 
  tibble(rep = 1:5000) |>
  mutate(
    estimates = map(rep, ~ boot_once(weather_df))
  ) |>
  unnest(estimates)
```
Distribution of r^2
```{r}
boot_results |>
  ggplot(aes(x = r2)) +
  geom_histogram(bins = 30) +
  labs(
    x = expression(r^2),
    y = "Count",
    title = expression("Bootstrap distribution of " * r^2)
  )
```
Distribution of beta 1 hat/beta 2 hat
```{r}
boot_results |>
  ggplot(aes(x = beta_ratio)) +
  geom_histogram(bins = 30) +
  labs(
    x = expression(hat(beta)[1] / hat(beta)[2]),
    y = "Count",
    title = expression("Bootstrap distribution of " * hat(beta)[1] / hat(beta)[2])
  )
```
Comment: 
The bootstrap distribution of r^2 is fairly concentrated, suggesting that model fit is reasonable stable across resamples. 
95% bootstrap Cls using quantiles 
```{r}
ci_summary = 
  boot_results |>
  summarize(
    r2_lo = quantile(r2, 0.025),
    r2_hi = quantile(r2, 0.975),
    ratio_lo = quantile(beta_ratio, 0.025),
    ratio_hi = quantile(beta_ratio, 0.975)
  )

ci_summary
```


